---
title: Exploratory Data Analysis, Week of 11/28/2022
author: Daniel Sussman
date: '2022-11-29'
slug: []
categories: []
tags: []
description: ~
toc: yes
authors: []
series: []
lastmod: '2022-11-29T22:19:58-05:00'
featuredVideo: ~
featuredImage: ~
---

Our main data sets that we are comparing are the baseball statistics and the payroll statistics. Our exploratory data analysis is continuing by looking for trends in these data sets combined, whereas before we looked at trends in each data set individually. We decided to no longer use the weather dataset.

```{r loading_data, echo=F}
suppressPackageStartupMessages(library(tidyverse))

load(here::here("dataset", "baseball.RData"))
load(here::here("dataset","mlb_payroll.Rdata"))
load(here::here("dataset","win_perc.Rdata"))
```

We first are looking at the basics, seeing if there is a correlation between winning and payroll data. To do this, we plot the payroll by the winning percentage, and facet by team. We then added a line of best fit to the data, which allows us to see if the overall trend is up, down, or zero for each team.

```{r trends_1,fig.height=20, echo=F}
combined<-full_join(mlb_payroll,win_perc,by=c('year','h_name'))

ggplot(combined,aes(log(payroll),win_perc))+geom_point()+
                                            geom_smooth(method='glm')+
                                            facet_wrap(vars(h_name),ncol=3)
```

Some teams clearly have an upward trend, others clearly downward, and some have no clear trend. From this graph, we chose teams to focus in on to see if there is anything else that contributes to the trend we see and could be correlated. We chose to look more closely at Pittsburgh (upward trend) and Chicago (downward trend).

We next looked at payroll and win percentage, but faceted by year instead of home team.

```{r trends_2,fig.height=20, echo=F}
ggplot(combined,aes(log(payroll),win_perc))+geom_point(aes(color=h_name))+
                                            facet_wrap(vars(year),ncol=3)+
                                            theme(legend.position = 'bottom')
```

While looking at each individual year appears to have no trend, when looking from year to year, we can see which teams continuously perform better and which teams continuously perform worse. For almost all of the years, the Yankees are outliers and typically pay the most money while also having a rather high win percentage. In 2014, the Dodgers surpass the Yankees in payroll, and also have a higher win percentage. In the later years, the Astros tend to have a low win percentage, however it can be seen that as they pay more money, this winning percentage improves and they are no longer an outlier on the lower end.These three teams are also worth looking into.

We have looked at payroll data versus number of homeruns, hits, double plays, doubles, and strikeouts. Correlation was only found in the number of strikeouts. We will continue to explore more variables to uncover all possible trends.