---
title: Data Loading and Cleaning, Week of 10/17/2022
author: Daniel Sussman
date: '2022-10-19'
slug: []
categories: []
tags: []
description: ~
toc: yes
authors: []
series: []
lastmod: '2022-10-19T09:38:50-04:00'
featuredVideo: ~
featuredImage: ~
---

TEST EDIT

Our data set of MLB game statistics is >50mb and can be found in the `dataset-ignore` folder. After reading in the CSV file, we converted it into a tibble. 

```{r}
source(
  here::here("static", "load_and_clean_data.R"),
  echo = TRUE
)

library(tidyverse)

data <- read_csv(here::here("dataset", "dataset-ignore", "game_logs.csv"))
(baseball <- as_tibble(data))
```

\n
\n

The date in the data is inserted as an integer, so using separate, we separated the date into three columns: `year`, `month`, and `day`. If we need to combine these, we can easily do so later on using lubridate.

```{r}
(baseball <- baseball %>% separate(date,into=c("year","month_day"),sep=4) %>% 
             separate(month_day,into=c("month","day"),sep=2))
```
             
\n
\n

Since this is a large dataset, we are considering working with those baseball games that took place between the years 2000 and 2016. Data on games from the 1970s and 1990s are missing so we chose to take a subset of more recent years of baseball. From the earliest years of the dataset, there are many variables with missing data as it might be too far in the past to recover such specific statistics of those games. Additionally, over the centuries of baseball, the rules of the game have changed so taking a subset of only the most recent years would help the data be more consistent.

```{r}
years<-c(2000,2001,2002,2003,2004,2005,2006,2007,2008,2009,2010,2011,2012,2013,2014,2015,2016)
(baseball <- baseball%>%filter(year %in% years))
```

\n
\n


This data set does not specify whether the home team wins or loses, but it does contain information on the home team's and the away team's number of runs. Since we are interested in seeing if a team wins or loses, we created a column in the dataset called `score_diff`, which takes the number of the visiting team's runs subtracted from the number of the one team's runs. A positive value represents a win, a negative value represents a loss, and a zero is a tie. The higher absolute values represent a bigger win/loss.

```{r}
(baseball <- baseball %>% mutate(score_diff=h_score-v_score))
```


\n
\n

Some filters we are considering are to filter by home team.

```{r}
baseball %>% filter(h_name=="BOS")
```

We filtered to select Boston Red Sox as the home team, since we have already found weather data for Boston and are considering taking this into account later on.

For our initial exploration of the data, we first looked at the average runs scored by each team at home as well as away arranged in descending order. Here, we can see which team scores the most runs on average. The Colorado Rockies score the most runs on average at home while the New York Yankees score the most runs on average away. Using our created variable `score_diff`, we can also observe which teams tend to lose or win more often. 

``` {r}
#Average runs scored by home teams
baseball %>% group_by(h_name) %>% summarize(average_score = mean(h_score)) %>% arrange(desc(average_score))

#Average runs scored by away/visiting teams
baseball %>% group_by(v_name) %>% summarize(average_score = mean(v_score)) %>% arrange(desc(average_score))

#Proportion of games won by home team
baseball %>% group_by(h_name) %>% summarize(win_prop = mean(score_diff>0)) %>% arrange(desc(win_prop))

#Proportion of games won by away/visiting team
baseball %>% group_by(v_name) %>% summarize(win_prop = mean(score_diff>0)) %>% arrange(desc(win_prop))
```

We can also look at `score_diff` for the Boston Red Sox on different days of the week. From the boxplot, it does not seem like the day of the week has an effect on whether the Red Sox win or lose. 

```{r}
baseball %>% filter(h_name == "BOS") %>% ggplot(aes(x = reorder(day_of_week, score_diff), y = score_diff)) + geom_boxplot()
```


