---
title: Initial Trends and Analysis, Week of 10/31/2022
author: Daniel Sussman
date: '2022-11-03'
slug: []
categories: []
tags: []
description: ~
toc: yes
authors: []
series: []
lastmod: '2022-11-03T19:02:24-04:00'
featuredVideo: ~
featuredImage: ~
---

Before analysis, the data must be loaded and mutated with extra columns we need for analysis.

```{r loading_data}
suppressPackageStartupMessages(library(tidyverse))
load(here::here("dataset", "baseball.RData"))
```

Graph #1 - Truly Exploratory Data: 

While exploring trends in the data, we set out to see if starting pitcher has an impact on the outcome. In the following code, we compare the starting pitchers in each season to both the proportion of their games played that they won, and the number of games that they won. We specifically are focusing on Boston Red Sox at their home stadium, but hope to explore more teams as well as away games in the future.

The first step is to create different data sets to see the number of games each pitcher wins, the number of games each pitcher plays, and the proportion of wins to the games played.

```{r graph1}
(pitcher_win<-baseball %>% filter(h_name=="BOS",h_score>v_score) %>% 
                           group_by(h_starting_pitcher_name,year) %>% 
                           count(h_starting_pitcher_name) %>% 
                           rename(games_won=n) %>% 
                           arrange(games_won))

(pitcher_played<-baseball %>% filter(h_name=="BOS") %>% 
                              group_by(h_starting_pitcher_name,year) %>% 
                              count(h_starting_pitcher_name) %>% 
                              rename(games_played=n) %>% 
                              arrange(games_played))

(pitcher_prop<-full_join(pitcher_played,pitcher_win) %>% 
               mutate(prop=games_won/games_played) %>% 
               replace_na(list(games_won=0,prop=0,games_played=0)) %>% 
               arrange(prop))
```
 
With these data sets, we can now graph both the proportion of games each pitcher wins and the number of games each pitcher wins. Each year corresponds to an individual season.

```{r graph1cont,fig.height=10,fig.width=10}

pitcher_prop %>% group_by(year) %>%
                 ggplot(aes(x=reorder(h_starting_pitcher_name,prop),y=prop,fill=h_starting_pitcher_name))+
                 geom_col(show.legend=FALSE)+
                 coord_flip()+
                 facet_wrap(vars(year),scales="free_y",ncol=3)+
                 labs(title="Proportion of Games Won for each Starting Pitcher of the Boston Red Sox")+
                 xlab("Proportion")+
                 ylab("Pitcher")

pitcher_win %>% group_by(year) %>%
                ggplot(aes(h_starting_pitcher_name,games_won,fill=h_starting_pitcher_name))+
                geom_col(show.legend=FALSE)+
                coord_flip()+
                facet_wrap(vars(year),scales="free_y",ncol=3)+
                labs(title="Games Won for each Starting Pitcher of the Boston Red Sox")+
                xlab("Number of Games Won")+
                ylab("Pitcher by ID")
```

There are clearly starting pitchers that win more games than others. In this case, the number of games won offers a little better insight than the proportion of games won, since multiple pitchers only started one game and won that one game, so are given a 100% win percentage. Starting pitchers is a factor that can definitely be explored further.

Graph #2 - An experiment in the accuracy of the Pythagorean Winning Percentage (PWP)

$$Pythagroean\ Winning\ Percentage = runs\ scored\ ^{2} / (run\ scored\ ^{2} + runs\ allowed\  ^{2})$$
 
```{r, graph2}
#win/loss column (1 = win, 0 = loss or tie)
h_win_loss <- function(h, v) {
  ifelse(h > v, 1, 0)
  
  # alternative to win loss function:
  # mutate(outcome = (ifelse(h_score>v_score, 1, 0))
}
(baseball <- baseball %>% mutate(h_win_loss = h_win_loss(h_score, v_score)))

#red sox games
bos <- baseball %>% filter(h_name == "BOS"|v_name == "BOS")

#for all teams
#total games won
home_wins <- baseball %>% group_by(h_name, year) %>% summarize(h_wins = sum(h_win_loss == 1))
away_wins <- baseball %>% group_by(v_name, year) %>% summarize(a_wins = sum(h_win_loss == 0))
team_year_wins <- home_wins %>% add_column(a_wins = away_wins$a_wins) %>% mutate(games_won = h_wins + a_wins)
#total games played
home_games <- baseball %>% group_by(h_name, year) %>% summarize(home_games = n())
away_games <- baseball %>% group_by(v_name, year) %>% summarize(away_games = n())
win_perc <- home_games %>% add_column(away_games = away_games$away_games) %>% mutate(games_played = home_games + away_games) %>% add_column(games_won = team_year_wins$games_won) %>% mutate(win_perc = games_won/games_played)

#pythagorean winning percentage
runs_h <- baseball %>% group_by(h_name, year) %>% summarize(runs_at_home = sum(h_score))
runs_a <- baseball %>% group_by(v_name, year) %>% summarize(runs_away = sum(v_score))
runs_total <- runs_h %>% add_column(runs_away = runs_a$runs_away) %>% mutate(runs_total = runs_away +runs_at_home)


h_allowed <- baseball %>% group_by(h_name, year) %>% summarize(home_allowed = sum(v_score))
a_allowed <- baseball %>% group_by(v_name, year) %>% summarize(away_allowed = sum(h_score))
allowed_total <- h_allowed %>% add_column(away_allowed = a_allowed$away_allowed) %>% mutate(allowed_total = home_allowed + away_allowed)

teams_pythag <- runs_total %>% add_column(allowed_total = allowed_total$allowed_total) %>% mutate(pythag_win = ((runs_total)^2)/(((runs_total)^2) + ((allowed_total)^2))) %>% add_column(win_perc = win_perc$win_perc)


#barplot for Boston Red Sox
teams_pythag %>% filter(h_name == "BOS") %>% 
  select(year, win_perc, pythag_win) %>% 
  pivot_longer(-(h_name:year), names_to = "pythag_actual", values_to = "percentage") %>%
  ggplot(aes(x=year, y=percentage, fill = pythag_actual)) + 
  geom_bar(stat = "identity", position = "dodge") + 
  scale_x_discrete(guide = guide_axis(n.dodge=2)) + 
          
           labs(title = "Pythagorean Winning Prediction versus Actual Winning Percentage", 
                x = "Years", 
                y = "Percentage", 
                color = "Percentage Calculation")
  labs(title = "Pythagorean Winnng Prediction versus Actual Winning Percentage", 
    x = "Years", 
    y = "Percentage", 
    color = "Percentage Calculation")

```
Because it can be hard to view how different these two values are, we can run a T-test and check the statistical significance of the prediction we've made. If we had hypothesized if this is an appropriate estimate for wins, we would accept or reject our hypothesis on the basis of the T-test: 

``` {r, graph2_cont}
# different representation:
bos <- teams_pythag %>% filter(h_name=="BOS")
(stat_sig <- t.test(bos$pythag_win, bos$win_perc, paired = TRUE))

```
Our p-value is greater than 0.05 (a value 0.05475) which indicated that our results are not statistically significant. To say that our prediction is not statistically significant from the actual winning percentage is to say the values are more or less similar and that this metric is a good predictor of wins. 

The next step after this data exploration would be to reproduce the result for more teams. 


Graph #3 - A logistical model predicting wins or losses

```{r}
baseballlog <- baseball %>% filter(h_name=="BOS")

baseballlog

mylogit <- glm(h_win_loss ~ h_errors + v_errors + h_left_on_base + v_left_on_base + h_homeruns + v_homeruns, data = baseballlog, family = "binomial")

summary(mylogit)

confint(mylogit)

meandata <- with(baseballlog, data.frame(h_errors = mean(h_errors), v_errors = mean(v_errors), h_left_on_base = mean(h_left_on_base), v_left_on_base = mean(v_left_on_base), h_homeruns = mean(h_homeruns), v_homeruns = mean(v_homeruns)))

meandata
```