---
title: Interactive Plot, Week of 11/14/22
author: Daniel Sussman
date: '2022-11-14'
slug: []
categories: []
tags: []
description: ~
toc: yes
authors: []
series: []
lastmod: '2022-11-14T09:43:26-05:00'
featuredVideo: ~
featuredImage: ~
---

For our interactive, we want to put the geographic location of all of the stadiums onto a map of the United States, colored by division. This was a good contribution to the website because it provides context and  When you click on a stadium, the team name and winning percentage of games played will pop up. Using the drop down menu, you can select one year for the winning percentages to be displayed from. 

The personalized aspect will be how the user has the ability to choose a stadium; they will be able to investigate the statistics for their favorite teams.


* Our rough outline of our interactive is shown below. We still need to convert to Shiny.
```{r, interactive}
library(tidyverse)
library(rvest)
library(tibble)
library(readxl)
library(leaflet)
library(dplyr)
library(readr)
library(googlesheets4)



load(here::here("dataset", "baseball.RData"))



load(here::here("dataset", "baseball.RData"))
stadiums <- read_sheet("https://docs.google.com/spreadsheets/d/1p0R5qqR7XjoRG2mR5E1D_trlygHSqMOUdMgMpzq0gjU/htmlview")


AL_east <- c('BAL', 'BOS', 'NYA', 'TBR', 'TOR')
AL_central <- c('MIN', 'CLE', 'CHW', 'DET','KCR')
AL_west <- c('HOU', 'SEA', 'OAK', 'TEX', 'LAA')
NL_east <- c('ATL', 'MIA', 'NYM', 'PHI', 'WAS')
NL_central<- c('CHC', 'CIN', 'MIL', 'PIT','STL')
NL_west <- c('ARI', 'COL', 'LAD', 'SFG', 'SDP')

mlbDivision <- function(x){
    
  if(x %in% AL_east){y <- 'AL East'} 
  else if(x %in% AL_central){y <- 'AL Central'} 
  else if(x %in% AL_west){y <- 'AL West'} 
  else if(x %in% NL_east){y <- 'NL East'} 
  else if(x %in% c(NL_central)){y <- 'NL Central'} 
  else { y <- 'NL West'}
  return(y)
}

icons <- makeAwesomeIcon(
    icon = 'circle',
    iconColor = 'black',
    library = 'glyphicon',
    markerColor = ifelse(stadiums$Abbreviation %in% AL_east, 'red',
                  ifelse(stadiums$`Team(s)` %in% AL_central,'blue',
                  ifelse(stadiums$`Team(s)` %in% AL_west, 'orange',
                  ifelse(stadiums$`Team(s)` %in% NL_east,'green',
                  ifelse(stadiums$`Team(s)` %in% NL_central,'darkpurple','black'))))))



mlbColor <- function(x){
    
  if(x == 'AL East'){y <- 'blue'}
  else if(x == 'AL Central'){y <- 'green'}
  else if(x == 'AL West'){y <- 'yellow'}
  else if(x == 'NL East'){y <- 'red'}
  else if(x == 'NL Central'){y <- 'purple'} 
  else {y <- 'black'}
  return(y)
  }

# 
# stadiums$Abbreviation[stadiums$Abbreviation == 'CHC'] <- 'CHN'
# stadiums$Abbreviation[stadiums$Abbreviation == 'CHW'] <- 'CHA'
# stadiums$Abbreviation[stadiums$Abbreviation == 'KCR'] <- 'KCN'
# stadiums$Abbreviation[stadiums$Abbreviation == 'LAA'] <- 'ANA'
# stadiums$Abbreviation[stadiums$Abbreviation == 'LAD'] <- 'LAN'
# stadiums$Abbreviation[stadiums$Abbreviation == 'MIA'] <- 'FLO'
# stadiums$Abbreviation[stadiums$Abbreviation == 'NYM'] <- 'NYN'
# stadiums$Abbreviation[stadiums$Abbreviation == 'NYY'] <- 'NYA'
# stadiums$Abbreviation[stadiums$Abbreviation == 'SDP'] <- 'SDN'
# stadiums$Abbreviation[stadiums$Abbreviation == 'SFG'] <- 'SFN'
# stadiums$Abbreviation[stadiums$Abbreviation == 'STL'] <- 'SLN'
# stadiums$Abbreviation[stadiums$Abbreviation == 'TBR'] <- 'TBA'

MLBstadiums <- stadiums %>% filter(Sport == "MLB") %>% 
  mutate(Division = sapply(X = Abbreviation, FUN = mlbDivision),
           DivisionColor = sapply(X = Division, FUN = mlbColor),
           teamLabs = paste0('Team: ',Abbreviation,'\nField: ',Venue))



baseball_w_stadiums <- merge(baseball, stadiums, by.x=c("h_name"), 
                                                 by.y =("Abbreviation")) 

  
makeColorsandNames <- data.frame(divisions = c('AL East','AL Central','AL West','NL East','NL Central','NL West'), division.cent = c('#CC0000','#3399FF','#FFA500','#9ACD32','#483D8B','#000000'))
  
  
leaflet(MLBstadiums) %>% 
      setView(lng = -98.5795, lat = 39.8283, zoom = 5) %>% 
      addTiles() %>% 
      addAwesomeMarkers(~Longitude, ~Latitude, icon = icons, label = ~as.character(Venue), popup = ~as.character(Sport)) %>% 
      addLegend(position = 'bottomleft', colors = makeColorsandNames[,2],labels = makeColorsandNames[,1],opacity = 1,title = 'Divisions')

    
```

```{r, combining sets}
load(here::here("dataset", "win_perc.RData"))
shiny_wins <- win_perc %>%  select(h_name, year, win_perc)
shiny_wins <- pivot_wider(shiny_wins, id_cols = h_name, names_from = year, values_from = win_perc)
```

```{r, shiny map}
#rsconnect::setAccountInfo(name='bu-ma415-hgb-stadium-map', token='AC86430056BEE712DADDBB47C5B56B5A', secret='sPdprlQo0fcDettSUrK/b++IIRdzexqNZs7jiSa3')

# library(rsconnect)
# rsconnect::deployApp('shiny_app/stadium_map_2/app.R')
# 
# library(shiny)
# 
# deployApp()
# runApp()
```


